
---
title: "Lecture_4_MACSS"
author: "Auffhammer"
date: "2024-09-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now let's get ambitious. We are going to generate some data on health status of some made up folks and play random sampling and random assignment. This forst exercise lets you simulate what happens if insurance is randomly assigned, versus if it is not.... It will also show us how to do the t-test to compare group means and how to use a canned routine to calculate these automatically, which is generally better (since the right test depends on the nature of the random variable you are looking at.)

```{r Health}
rm(list = ls()) # clear memory
library(dplyr)
library(MASS)
library(crosstable)
library(flextable)
library(ggplot2)
library(plyr)
set.seed(22092008) # set random number generator seed
n <- 1000 # Sample Size
mu <- c(0, 0,0)
# a <- 0.5 #Gender Income Covariance
# b <- 0.1 #Gender Insurance Covariance
# c <- 0.8 #Income Insurance

# If insurance were randomized
a <- 0.5 # Set to 0.5 as default
b <- 0.0 # Set to 0.1 as default
c <- 0.0 # Set to 0.8 as default

# Some betas for later
b1 <-1 #Gender Beta
b2 <-5 # Income Beta
b3 <-3 #Insurance
shifter <- 30
Sigma <- matrix(c(1, a, b, a, 1, c,b, c, 1), nrow=3)
data = mvrnorm(n, mu, Sigma, empirical=FALSE)
Gender = data[, 1]  # standard normal (mu=0, sd=1)
Income = data[, 2]  # standard normal (mu=0, sd=1)
Insurance= data[, 3]  # standard normal (mu=0, sd=1)

# Gender and income should be binary
Ins <- Insurance>0
Gend <- Gender>0
cor(Ins,Gend) 
cor(Ins,Income) 
cor(Gend,Income)

# We are going to generate some arbitrary Health Index
Health <- shifter +  rnorm(n,mean=0,sd=1) + b1*Gend + b2*Income + b3* Ins

# Let's do a manual comparison of Health Across the Insured and Uninsured.
mydata <- data.frame(Income, Gend, Health, Ins)
# Calculate Means by Group (using ddply). 
mu <- ddply(mydata, "Ins", summarise, grp.mean=mean(Health))

# Plot my health outcome by Insurance Status in Pretty Graph
ggplot(mydata, aes(x=Health, color=Ins, fill=Ins)) +
  scale_color_manual(values=c("#002676", "#FC9313")) +
  scale_fill_manual(values=c("#002676", "#FC9313")) +
  geom_histogram(alpha=0.1, position="identity", bins=50)+
  geom_vline(data=mu, aes(xintercept=grp.mean, color=Ins),
             linetype="dashed") +
  labs(title="Health Outcomes by Insurance Status",x="Health Index", y = "Count")+
  theme_classic()

# Let's compare across treatment
# First - do this by hand. Difference in means. Unknown and uneuqal variances.

M1 <- mean(mydata[Ins == 'TRUE', 'Health']) 
M2 <- mean(mydata[Ins == 'FALSE', 'Health'])
n1 <- sum(Ins)
n2 <- n-n1
V1 <- var(mydata[Ins == 'TRUE', 'Health']) 
V2 <- var(mydata[Ins == 'FALSE', 'Health']) 
S <- sqrt((V1 / n1) + (V2 / n2))
statistic <- (M1 - M2 - 0) / S
print(statistic)

t_health <- t.test(Health ~ Ins)
print(t_health)
t_inc <- t.test(Income ~ Ins)
print(t_inc)
my_test_args=crosstable_test_args(show_method=FALSE)
ft1 <- crosstable(mydata,by="Ins", test=TRUE, funs=c(mean=mean),test_args=my_test_args) %>% 
  as_flextable()
 print (ft1)

```

Now let's turn to some simple regression analysis. It is so simple, my teenager can do it. In fact, I checked and he can. That said, interpreting what it tells you is going to be the art form. 

```{r regressions}
rm(list = ls()) # clear memory
library(dplyr)
library(MASS)
library(stargazer) # For pretty Regression Tables
setwd("/Users/auffhammer/Library/CloudStorage/Dropbox/06_Teaching/MACSS/2024/code/public-repository-1/week_4")

airfares <- read.csv("airfares.csv")
#Let's plot some data. 
ggplot(airfares, aes(x=dist, y=fare)) + 
  geom_point(alpha=0.5, shape=16, fill="#002676", color="#002676", size=2)+
  geom_smooth(method=lm, color="#FC9313")+
  labs(title="Airfare by Distance",
       x="Distance (Miles)", y = "Fare (US$)")+
  theme_classic()  


#Let's run a regression.  
planes <- lm(airfares$fare ~ airfares$dist)
# Let's record some residuals and join them to our data frame. 
airfares$res <- planes$resid

#Let's make some nice looking regression output. 

stargazer(planes, type='text', digits = 3, title = 'Linear Airfare Distance Regression', style = 'qje')

# Plot Residuals - Playing with colors (HEX Colors - official Cal!
# Also meesing with background and Axis Labels. )
ggplot(airfares, aes(x=dist, y=res)) + 
  geom_point(alpha=0.5, shape=16, fill="#002676", color="#002676", size=2)+
  geom_smooth(method=lm, se=FALSE, color="#FC9313")+
  labs(title="Airfare by Distance",
       x="Distance (Miles)", y = "Residuals")+
  theme_classic()  

# Much nicer than the junk I showed in lecture. Apologies. 


```

```{r avocados}
library(dplyr)
library(MASS)
library(stargazer) # For pretty Regression Tables
setwd("/Users/auffhammer/Library/CloudStorage/Dropbox/06_Teaching/MACSS/2024/code/public-repository-1/week_4")

avocado <- read.csv("avocado.csv")

#Let's plot some data. I am fitting a smoother (loess) to the data to see what the functional # form looks like. A bit more on this later. 

ggplot(avocado, aes(x=price_reg, y=quantity_reg)) + 
  geom_point(alpha=0.5, shape=16, fill="#002676", color="#002676", size=2)+
  geom_smooth(method=loess, color="#FC9313")+
  labs(title="Avocado Demand (Conventional)",
       x="Quantity (some units)", y = "Price (US$ per some unit)")+
  theme_classic()  

# Now run a regression of the linear model No transformation. 
avo_lin <- lm(avocado$quantity_reg ~ avocado$price_reg)

# Let's record some residuals and join them to our data frame. 
avocado$res <- avo_lin$resid

#Let's make some nice looking regression output. 
stargazer(avo_lin, type='text', digits = 3, title = 'Linear Price Regression', style = 'qje')

# Plot Residuals - same pretty graph as before. 
ggplot(avocado, aes(x=price_reg, y=res)) + 
  geom_point(alpha=0.5, shape=16, fill="#002676", color="#002676", size=2)+
  geom_smooth(method=lm, se=FALSE, color="#FC9313")+
  labs(title="Residual Plot - Avocado Demand",
       x="Price (US$ per something)", y = "Residuals")+
  theme_classic()  

# Transform our variables using natural logs. 
avocado$l_price <- log(avocado$price_reg)
avocado$l_q <- log(avocado$quantity_reg)

# Run the log log regression. 
avo_log <- lm(avocado$l_q ~ avocado$l_price)

# Let's record some residuals and join them to our data frame. 
avocado$resl <- avo_log$resid

# Plot Residuals - same pretty graph as before. 
ggplot(avocado, aes(x=l_price, y=resl)) + 
  geom_point(alpha=0.5, shape=16, fill="#002676", color="#002676", size=2)+
  geom_smooth(method=lm, se=FALSE, color="#FC9313")+
  labs(title="Residual Plot - Log-Log Avocado Demand Residuals",
       x="Log(Price) (US$ per something)", y = "Residuals")+
  theme_classic()  

```

